---
title: "Analyses Pilot"
author: "Jeremy Zehr"
date: "11/12/2019"
output: html_document
---

```{r setup, include=FALSE}
library(dplyr)
library(lme4)
library(ggplot2)

read.pcibex <- function(filepath, auto.colnames=TRUE, fun.col=function(col,cols){cols[cols==col]<-paste(col,"Ibex",sep=".");return(cols)}) {
  n.cols <- max(count.fields(filepath,sep=",",quote=NULL),na.rm=TRUE)
  if (auto.colnames){
    cols <- c()
    con <- file(filepath, "r")
    while ( TRUE ) {
      line <- readLines(con, n = 1, warn=FALSE)
      if ( length(line) == 0) {
        break
      }
      m <- regmatches(line,regexec("^# (\\d+)\\. (.+)\\.$",line))[[1]]
      if (length(m) == 3) {
        index <- as.numeric(m[2])
        value <- m[3]
        if (index < length(cols)){
          cols <- c()
        }
        if (is.function(fun.col)){
          cols <- fun.col(value,cols)
        }
        cols[index] <- value
        if (index == n.cols){
          break
        }
      }
    }
    close(con)
    return(read.csv(filepath, comment.char="#", header=FALSE, col.names=cols))
  }
  else{
    return(read.csv(filepath, comment.char="#", header=FALSE, col.names=seq(1:n.cols)))
  }
}
```

## Preparations and General observations

Reading the table, and making sure EventTime is numeric (important column)

```{r reading}
df.all <- read.pcibex("results40ppts.csv")

df.all$EventTime <- as.numeric(as.character(df.all$EventTime))
```

How long did each participant take?
```{r completion times}
df.all %>% group_by(id) %>% summarise(duration = (max(EventTime, na.rm=TRUE) - min(EventTime, na.rm=TRUE))/60000)
```

Preparing df.trials
```{r df.trials}
df.trials <- subset(df.all, grepl("input|practice", Type) & PennElementType%in%c("Scale","Var"))

# Forgot to log Dimension, adding it based on Item number:
dim.temp <- data.frame(Item.number=c(4:7,27:28), Dimension="Temperature", Nonce=FALSE)
dim.light <- data.frame(Item.number=c(8:11,29:30), Dimension="Light", Nonce=FALSE)
dim.sound <- data.frame(Item.number=c(12:15,31:32), Dimension="Sound", Nonce=TRUE)
dim.angle <- data.frame(Item.number=c(16:19,33:34), Dimension="Angle", Nonce=TRUE)
df.dim <- rbind(dim.temp,dim.light,dim.sound,dim.angle)
df.trials <- merge(df.trials,df.dim)
```

Focusing on input (leaving practice aside)
```{r df.input}
df.input <- df.trials %>% 
  filter(grepl("input", Type) & PennElementType=="Scale") %>%
  group_by(id, Dimension, Form, Force) %>%
  select(id,EventTime,Value,Dimension,Form,Force,Tested,Level,Nonce) %>%
  arrange(EventTime)
```

Did participants change direction while setting the dial?
```{r Hesitation}
serialdifference <- function(t){
  t<-as.numeric(as.character(t))
  return(t-c(t[1],t[1:(length(t)-1)]))
}

df.input <- df.input %>% mutate(ValueDiff=serialdifference(Value)) %>% mutate(ConstantChange=all(ValueDiff<=0)|all(ValueDiff>=0))

summary((df.input %>% group_by(id) %>% summarise(overall=mean(ConstantChange)))$overall)
df.input %>% group_by(Dimension) %>% summarise(overall=mean(ConstantChange)) %>% arrange(overall)
df.input %>% group_by(Dimension, Form, Force) %>% summarise(overall=mean(ConstantChange)) %>% arrange(overall)
```

Final values on trials
```{r finalvalues}
df.input.final <- df.input %>% filter(EventTime == max(EventTime))
df.input.final$Value <- as.numeric(as.character(df.input.final$Value))

df.input.final$Increase <- df.input.final$Value > df.input.final$Level
df.input.final$Decrease <- df.input.final$Value < df.input.final$Level
# AdjustedIncrease == !Increase for mid-light (dim) and mid-angle; else == Increase
df.input.final$AdjustedIncrease <- df.input.final$Increase
midsoundangle <- df.input.final$Force=="Mid"&(df.input.final$Dimension=="Light"|df.input.final$Dimension=="Angle")
df.input.final$AdjustedIncrease[midsoundangle] <- (df.input.final$Increase[midsoundangle]|df.input.final$Decrease[midsoundangle])&!df.input.final$Increase[midsoundangle]
```

For each adjective (and each participant), IncreaseDiff = 
    1     if    Increase on comparative  &  Decrease on excessive
   -1     if    Decrease on comparative  &  Increase on excessive
    0     if    Same on comparative  &  excessive
```{r increasediff}
df.input.final <- df.input.final %>% group_by(id,Dimension,Force) %>% arrange(Form) %>%                                 mutate(IncreaseDiff=rep(0-diff(Increase),2),AdjIncreaseDiff=rep(0-diff(AdjustedIncrease),2))
```

## Stats

Did every participant change the dial's value on every trial?
```{r inertia}
all(df.input.final$Increase|df.input.final$Decrease)
```

How many participants increased in each conditions?
```{r}
df.input.final %>% 
  group_by(Dimension, Form, Force) %>%
  summarise( NIncrease=sum(Increase) , N = length(Increase) )
```

Logistic model predicting (adjusted) increase from Force (mid v max) and Form (comp v exc)
```{r model}
df.input.final$id <- factor(df.input.final$id)
df.input.final$Form <- relevel(df.input.final$Form, "excessive")
df.input.final$Force <- relevel(df.input.final$Force, "Mid")
# Adding random structure for id wouldn't converge
md.all <- glmer(AdjustedIncrease ~ Form * Force + (1|Dimension), 
                family="binomial",control=glmerControl(optimizer="bobyqa"), data=df.input.final)
summary(md.all)
```
Global stats support the truly-central hypothesis

## Plots

General plot, by force (mid v max) and by form (comp v exc)
```{r generalplot}
ggplot(df.input.final %>% group_by(Force, Form) %>% summarise(AdjIncreaseProp = sum(AdjustedIncrease)/length(AdjustedIncrease), AdjIncreaseDiffProp = sum(AdjIncreaseDiff)/length(AdjIncreaseDiff))) +
  geom_bar(aes(x=Force, y=AdjIncreaseProp, fill=Form), stat="identity", position="dodge") +
  geom_bar(aes(x=Force,y=AdjIncreaseDiffProp), width=0.1, stat="identity", position="dodge")
```
Global picture is clear for Max (control) but complex for Mid

Real (FALSE) v Nonce (TRUE) adjectives
```{r plotnonce}
ggplot(df.input.final %>% group_by(Force, Form, Nonce) %>% summarise(AdjIncreaseProp = sum(AdjustedIncrease)/length(AdjustedIncrease), AdjIncreaseDiffProp = sum(AdjIncreaseDiff)/length(AdjIncreaseDiff))) +
  geom_bar(aes(x=Force,y=AdjIncreaseProp,fill=Form), stat="identity", position="dodge") +
  geom_bar(aes(x=Force,y=AdjIncreaseDiffProp), width=0.1, stat="identity", position="dodge") +
  facet_wrap(~Nonce)
```
Clear picture for real adjectives, complex for nonce adjectives

Plotting each adjective separately
```{r plotsperadjective}
# Raw Increase
ggplot(df.input.final %>% group_by(Force, Form, Dimension) %>% summarise(IncreaseProp = sum(Increase)/length(Increase), IncreaseDiffProp = mean(IncreaseDiff))) +
       geom_bar(aes(x=Force, y=IncreaseProp, fill=Form), stat="identity", position="dodge") +
       geom_bar(aes(x=Force,y=IncreaseDiffProp), width=0.1, stat="identity", position="dodge") +
        facet_wrap(~Dimension)
# Adjusted Increase
ggplot(df.input.final %>% group_by(Force, Form, Dimension) %>% summarise(AdjIncreaseProp = sum(AdjustedIncrease)/length(AdjustedIncrease), AdjIncreaseDiffProp = mean(AdjIncreaseDiff))) +
  geom_bar(aes(x=Force, y=AdjIncreaseProp, fill=Form), stat="identity", position="dodge") +
  geom_bar(aes(x=Force,y=AdjIncreaseDiffProp), width=0.1, stat="identity", position="dodge") +
  facet_wrap(~Dimension)
```

Did turning the sound off make a difference?
```{r plotsoundoff}
ggplot(merge(df.input.final, df.all %>% filter(PennElementName=="sound") %>% select(id, Value) %>% rename(Sound = Value)) %>% filter(Sound=="off") %>% group_by(Force, Form, Dimension) %>% summarise(AdjIncreaseProp = sum(AdjustedIncrease)/length(AdjustedIncrease), AdjIncreaseDiffProp = mean(AdjIncreaseDiff))) +
  geom_bar(aes(x=Force, y=AdjIncreaseProp, fill=Form), stat="identity", position="dodge") +
  geom_bar(aes(x=Force,y=AdjIncreaseDiffProp), width=0.1, stat="identity", position="dodge") +
  facet_wrap(~Dimension)
```
Apparently not!